---
title: "P8105 Homework 2"
author: "sarah_8105"
output: github_document
---

This is my second homework assignment for P8105.

```{r libraries}
library(tidyverse)
library(readxl)
library(readr)
```


## Problem 1

The Mr. Trash Wheel data set represents trash that has been removed from the Inner Harbor in Baltimore, Maryland by the Mr. Trash Wheel vessel. The data set contains information on the amount of trash collected by each dumpster by month and year, breaking down specific types of trash items. In this first portion of the exercise, I will read and clean the Mr. Trash Wheel data.

In this code chunk, I read in the Mr. Trash Wheel sheet, clean variable names, omit non-data entries and rows that do not include dumpster-specific data, and change the number of sports balls variable to an integer.

```{r}
trash_wheel_df = 
  read_excel(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

In the next portion of the exercise, I read in the monthly precipitation data for 2018 and 2017 from the Mr. Trash Wheel data. In this step, I omit rows without precipitation data and create a year variable.

```{r precip_import}
precipitation_18 = 
  read_excel(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation",
    skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precipitation_17 = 
  read_excel(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation",
    skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

In the next code chunk I combine the annual precipitation over 2018 and 2017 and create a month character variable.

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precipitation_df =
  bind_rows(precipitation_18, precipitation_17)

left_join(precipitation_df, month_df, by = "month")

```

The Mr. Trash Wheel data set contains `r nrow(trash_wheel_df)` observations and `r ncol(trash_wheel_df)` variables. 

The resulting precipitation data set contains `r nrow(precipitation_df)` observations and `r ncol(precipitation_df)` variables. 


 
## Problem 2

The next problem explores the NYC Transit data on subway station entrances and exits.

In the following code chunk, I read in the NYC Transit data, retaining:

* line
* station name
* latitude / longitude
* routes served
* entry
* vending 
* entrance type
* ADA compliance

I also convert entry from a character variable (_YES_ vs. _NO_) to a logical variable (_YES = TRUE_ vs. _NO = FALSE_).

```{r transit_import}
nyc_transit = read_csv(
  "./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = ifelse(entry == "yes", TRUE, FALSE))
nyc_transit
```

The resulting data set contains `r nrow(nyc_transit)` observations and `r ncol(nyc_transit)` variables. 
