---
title: "P8105 Homework 2"
author: "sarah_8105"
output: github_document
---

This is my second homework assignment for P8105.

```{r libraries}
library(tidyverse)
library(readxl)
```


## Problem 1

The Mr. Trash Wheel data set contains information on trash removal from the Inner Harbor in Baltimore, Maryland. In this first portion of the exercise, I will read and clean the Mr. Trash Wheel precipitation data for 2017 and 2018 and will conduct exploratory analyses of the data.

In this first code chunk, I read in the Mr. Trash Wheel sheet, clean variable names, and omit non-data entries and rows that do not include dumpster-specific data.

```{r}
trash_wheel = read_excel(
  "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = "A2:N406") %>%
  janitor::clean_names() %>%
  drop_na(dumpster)
trash_wheel
```


```{r precip_import}
precipitation_18 = read_excel(
  "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "2018 Precipitation",
  range = "A2:B14")
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.